Strong lensing selection effects

### 2022-07-22 paper/seleff.tex
q distribution

We described the surface mass density distribution of each galaxy as an elliptical de Vaucouleurs profile (a S\'{e}rsic profile with $n=4$).
Given the observed stellar mass of a galaxy, we assigned a half-mass radius by drawing it from the following distribution in $\log{\reff}$:
\begin{equation}\label{eq:masssize}
\pr(\log{\reff}) \sim \mathcal{N}(1.20 + 0.63(\log{\mobs} - 11.4), 0.14^2).
\end{equation}
Then, we assigned an axis ratio $q$ by drawing it from the following distribution:
\begin{equation}\label{eq:qdist}
\pr(q) \sim \,{\rm boh}.
\end{equation}
%The choice for the distributions of \Eref{eq:masssize} and \Eref{eq:qdist}
The choice for these distributions was motivated by observations of a sample of early-type galaxies. In Appendix~\ref{sect:appendixa} we explain how this sample was defined and how the coefficients of \Eref{eq:masssize} and \Eref{eq:qdist} were determined.

@@

To assign half-light radii and ellipticities to the simulated lenses we relied on observations of a sample of early-type galaxies selected from the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00}.
This sample was selected as follows. First we defined a narrow redshift slice around $z=0.2$. Then we applied a selection in colour, by choosing objects with $g-r>1.2$, and on S\'{e}rsic index, by selecting only galaxies with $n>2.5$. We used the S\'{e}rsic fit measurements by \citet{Mee++15} for this purpose. These cuts produced a sample of $8078$ galaxies.
%We focused on the $r-$band de Vaucouleurs profile measurements obtained by \citet{Mee++15}, for these objects.
%We then took measurements of their total flux, half-light radius and axis ratio from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}. Using the 
%We then took measurements of the total flux, half-light radius and axis ratio obtained by \citet{Mee++15} from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}. Using the 
%For each object, we took the measurement of its de Vaucouleurs model-based $r-band$ total flux, half-light radius and axis ratio, as provided by \citet{Mee++15}.
We then focused on the $r-$band de Vaucouleurs model-based photometric measurements of \citet{Mee++15}.
Using the $r-$band total flux from the de Vaucouleurs model and the stellar mass-to-light ratio estimates of \citet{Men++14}, we obtained measurements of $\mobs$. Finally, we fitted the stellar mass-size relation with the following model:
\begin{equation}
\log{\reff} \sim \mathcal{N}(\mu_{R,0} + \beta_R(\log{\mobs} - 11.4), \sigma_R^2).
\end{equation}
We obtained

>>>>

We described the surface mass density distribution of each galaxy as an elliptical de Vaucouleurs profile (a S\'{e}rsic profile with $n=4$).
Given the observed stellar mass of a galaxy, we assigned a half-mass radius by drawing it from the following distribution in $\log{\reff}$:
\begin{equation}\label{eq:masssize}
\pr(\log{\reff}) \sim \mathcal{N}(1.20 + 0.63(\log{\mobs} - 11.4), 0.14^2).
\end{equation}
Then, we assigned an axis ratio $q$ by drawing it from the following beta distribution:
\begin{equation}\label{eq:qdist}
\pr(q) \propto q^{\alpha-1}(1 - q)^{\beta-1},
\end{equation}
with $\alpha=6.28$ and $\beta=2.05$.
The choice for these distributions was motivated by observations of a sample of early-type galaxies. In Appendix~\ref{sect:appendixa} we explain how this sample was defined and how the coefficients of \Eref{eq:masssize} and \Eref{eq:qdist} were determined.

@@

To assign half-light radii and ellipticities to the simulated lenses we relied on observations of a sample of early-type galaxies selected from the Sloan Digital Sky Survey \citep[SDSS][]{Yor++00}.
This sample was selected as follows. First we defined a narrow redshift slice around $z=0.2$. Then we applied a selection in colour, by choosing objects with $g-r>1.2$, and on S\'{e}rsic index, by selecting only galaxies with $n>2.5$. We used the S\'{e}rsic fit measurements by \citet{Mee++15} for this purpose. These cuts produced a sample of $8078$ galaxies.
%We focused on the $r-$band de Vaucouleurs profile measurements obtained by \citet{Mee++15}, for these objects.
%We then took measurements of their total flux, half-light radius and axis ratio from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}. Using the 
%We then took measurements of the total flux, half-light radius and axis ratio obtained by \citet{Mee++15} from the de Vaucouleurs profile fits to the SDSS $r$-band data obtained by \citet{Mee++15}. Using the 
%For each object, we took the measurement of its de Vaucouleurs model-based $r-band$ total flux, half-light radius and axis ratio, as provided by \citet{Mee++15}.
We then focused on the $r-$band de Vaucouleurs model-based photometric measurements of \citet{Mee++15}.
Using the $r-$band total flux from the de Vaucouleurs model and the stellar mass-to-light ratio estimates of \citet{Men++14}, we obtained measurements of $\mobs$. Finally, we fitted the stellar mass-size relation with the following model:
\begin{equation}
\log{\reff} \sim \mathcal{N}(\mu_{R,0} + \beta_R(\log{\mobs} - 11.4), \sigma_R^2).
\end{equation}
We obtained $\mu_{R,0}=1.20$, $\beta_R=0.63$ and $\sigma_R=0.14$.

We then proceeded to fit for the axis ratio distribution of the same sample of galaxies.
\Fref{fig:qhist} shows a histogram of the observed distribution.
We fitted this with a beta disribution:
\begin{equation}\label{eq:betaappendix}
\pr(q) \propto q^{\alpha-1}(1 - q)^{\beta-1}.
\end{equation}
We obtained $\alpha=6.28$ and $\beta=2.05$.
\begin{figure}
\includegraphics[width=\columnwidth]{q_dist.eps}
\caption{
Distribution in axis ratio of a sample of early-type galaxies.
The sample consists of $8078$ galaxies from the SDSS, selected by means of cuts in redshift, colour and S\'{e}rsic index as explained in the text.
Measurements of the axis ratio are taken from the de Vaucouleurs model fits of \citet{Mee++15}.
The model curve is a beta distribution, \Eref{eq:betaappendix}, with $\alpha=6.28$ and $\beta=2.05$.
}
\end{figure}
